Pricing Games in Heterogeneous Parallel Networks

### 2025-02-10 thesis/slides.tex
Thesis: Pres: Final fixes

\begin{frame}{Preliminaries | Selfish routing (1\textsuperscript{st} level)}
	\begin{definition}
		For a given set of tolls $t$, a flow $x$ is a \textit{Nash equilibrium for $t$} if $\forall p \in [0, 1]$ and $\forall i \in N$ it holds that
		\begin{equation*}
			c_{f(p)}(p) \leq c_i(p)
		\end{equation*}
	\end{definition}
	N.E. always exists and is unique w.r.t. costs for homogeneous\footcite{beckmann1956studies} and heterogeneous\footcites{1973JSP.....7..295S}{MILCHTAICH1996111} networks.
	We denote it by $x(t)$.
\end{frame}

@@

\begin{frame}{Preliminaries | Game equivalence}
	\begin{definition}[Game equivalence]
		Let $\Gm_1, \Gm_2$ be two $n$-link Network Congestion Games.
		$\Gm_1$ and $\Gm_2$ are called \textbf{equivalent} if and only if for all tolls $t \in \R_+^{|N|}$ it holds that $x^{(1)}(t) = x^{(2)}(t)$, with $x^{(1)}(t), x^{(2)}(t)$ being the Nash Equilibria for $t$ in $\Gm_1, \Gm_2$ respectively.
	\end{definition}
	Equivalency means same tolls cause same flow.\pause

@@

\begin{frame}{Sensitivity split | Properties}
	\begin{center}
		\begin{multicols}{2}
			% Left Column: Network Graph
			\begin{tikzpicture}[scale=1.2, baseline=40pt, every node/.style={font=\small}]
				% Nodes
				\node[circle, draw, fill=blue!20, minimum size=0.8cm] (s) at (0,0) {$s$};
				\node[circle, draw, fill=blue!20, minimum size=0.8cm] (t) at (4,0) {$t$};

@@

				% Description
				\node[align=center] at (0.7, -0.8) {$\alpha(p) = p + 1$};
			\end{tikzpicture}
		\end{multicols}
	\end{center}
	\[x(0) = \left(\frac23, \frac13\right) \qquad \alpha(x_1(0)) = \frac53 \quad \alpha(x_2(0)) = \frac43\]
\end{frame}

\begin{frame}{Sensitivity split | Properties}
	\begin{figure}
		\centering
		\begin{tikzpicture}[x=2cm, y=2cm, every node/.style={font=\small}]
			% Axes
			\draw[->] (0,0) -- (-1.1,0);
			\draw[->] (0,0) -- (2.1,0) node[right] {$t_1 - t_2$};
			\draw[->] (0,0) -- (0,1.2) node[above] {$x_1(t)$};

@@

\begin{frame}{Sensitivity split | Properties}
	\begin{figure}
		\centering
		\begin{tikzpicture}[x=2cm, y=2cm, every node/.style={font=\small}]
			% Axes
			\draw[->] (0,0) -- (-1.1,0);
			\draw[->] (0,0) -- (2.1,0) node[right] {$t_1 - t_2$};
			\draw[->] (0,0) -- (0,2.2) node[above] {$\as(t)$};

@@

	Assume $\Gm^\prime = (N, \ell, 1)$ and we calculate
	\[t^{*\prime} = \left(\frac53, \frac43\right) \qquad x(t^*) = \left(\frac59, \frac49\right)\]\pause

@@

\begin{frame}{Step distribution functions | Pricing N.E. Algorithm}
	\textbf{Summary:} Given $\Gm = ([2], \ell, \alpha), \alpha = (\alpha_k)_{k \in [m]}$
	\begin{enumerate}
		\item Consider $\Gm^\prime = ([2], \ell, 1)$ and calculate $t^{*\prime}, x^\prime(t^{*\prime})$.\pause
		\item Calculate $\as(t^*)$ as the $\alpha$ value of the lower split in $x^\prime(t^{*\prime})$.\pause
		\item Calculate the candidate $t^* = \tfrac{t^{*\prime}}{\as(t^*)}$ and $\Pi_1(t^*), \Pi_2(t^*)$.\pause
		\item For each $\alpha_k \ne \as(t^*)$, and for each $i \in [2]$:\pause
		\begin{enumerate}[$4.1$]
			\item Consider $\Gm_k = ([2], \ell, \alpha_k)$ and calculate $B_{i, k}(t_{-i}^*)$ and tolls $t_{i, k}$ for which $\as(t_i, t_{-i}^*) = \alpha_k$, resulting in two toll ranges for $t_i$.\pause
			\item Search for $\Pi_{i, k}$, the max profit for $i$ within those $t_i$ ranges:\pause
			\begin{enumerate}[$4.2.1$]
				\item If $B_{i, k}(t_{-i}^*)$ is within the $t_i$ ranges, $\Pi_{i, k} = \Pi_i(B_{i, k}(t_{-i}^*), t_{-i}^*)$.\pause
				\item Otherwise, $\Pi_{i, k}$ is max profit among the ranges' edge tolls ($4$).
			\end{enumerate}\pause
			\item If $\Pi_i < \Pi_{i, k}$, then $\Gm$ has NO pricing Nash Equilibrium.
		\end{enumerate}\pause
		\item If all of the above checks fail, return $t^*$.
	\end{enumerate}\pause
	\textbf{Complexity:} $\mathcal{O}(m)$
\end{frame}

@@

\begin{frame}{Future work}
	\textbf{Pseudo-heterogeneous games}\pause
	\begin{itemize}
		\item Connection between similar games with different fixed distribution functions\pause
		\item What happens to flow, tolls, profit when fixed $t_{-i}$?\pause
		\item (Conjectures in the Appendix)
	\end{itemize}\pause

@@

\begin{frame}{Future work}
	\textbf{Continuous distribution functions}\pause
	\begin{itemize}
		\item No gaps!\pause
		\item Function series approximation ($m \rightarrow +\infty$)\pause
		\item Linear $\alpha$ maintains the linear system\pause
		\item Strange behaviour when $0$ is an accumulation point in $\alpha$
	\end{itemize}\pause

>>>>

\begin{frame}{Preliminaries | Selfish routing (1\textsuperscript{st} level)}
	\begin{definition}
		For a given set of tolls $t$, a flow $x$ is a \textit{Nash equilibrium for $t$} if $\forall p \in [0, 1]$ and $\forall i \in N$ it holds that
		\begin{equation*}
			c_{f(p)}(p) \leq c_i(p)
		\end{equation*}
	\end{definition}\pause
	N.E. always exists and is unique w.r.t. costs for homogeneous\footcite{beckmann1956studies} and heterogeneous\footcites{1973JSP.....7..295S}{MILCHTAICH1996111} networks.
	We denote it by $x(t)$.
\end{frame}

@@

\begin{frame}{Preliminaries | Game equivalence}
	\begin{definition}[Game equivalence]
		Let $\Gm_1, \Gm_2$ be two $n$-link Network Congestion Games.
		$\Gm_1$ and $\Gm_2$ are called \textbf{equivalent} if and only if for all tolls $t \in \R_+^n$ it holds that $x^{(1)}(t) = x^{(2)}(t)$, with $x^{(1)}(t), x^{(2)}(t)$ being the Nash Equilibria for $t$ in $\Gm_1, \Gm_2$ respectively.
	\end{definition}
	Equivalency means same tolls cause same flow.\pause

@@

\begin{frame}{Sensitivity split | Example}
	\begin{center}
		\begin{multicols}{2}
			% Left Column: Network Graph
			\begin{tikzpicture}[scale=1.2, baseline=40pt, every node/.style={font=\small}]
				% Nodes
				\node[circle, draw, fill=blue!20, minimum size=0.8cm] (s) at (0,0) {$s$};
				\node[circle, draw, fill=blue!20, minimum size=0.8cm] (t) at (4,0) {$t$};

@@

				% Description
				\node[align=center] at (0.7, -0.8) {$\alpha(p) = p + 1$};
			\end{tikzpicture}
		\end{multicols}
	\end{center}
	\[x(0) = \left(\frac23, \frac13\right) \qquad \as(t) = \alpha(\text{lower split flow})\]
\end{frame}

\begin{frame}{Sensitivity split | Example}
	\begin{figure}
		\centering
		\begin{tikzpicture}[x=2cm, y=2cm, every node/.style={font=\small}]
			% Axes
			\draw[->] (0,0) -- (-1.1,0);
			\draw[->] (0,0) -- (2.1,0) node[right] {$t_1 - t_2$};
			\draw[->] (0,0) -- (0,1.2) node[above] {$x_1(t)$};

@@

\begin{frame}{Sensitivity split | Example}
	\begin{figure}
		\centering
		\begin{tikzpicture}[x=2cm, y=2cm, every node/.style={font=\small}]
			% Axes
			\draw[->] (0,0) -- (-1.1,0);
			\draw[->] (0,0) -- (2.1,0) node[right] {$t_1 - t_2$};
			\draw[->] (0,0) -- (0,2.2) node[above] {$\as(t)$};

@@

	Assume $\Gm^\prime = (N, \ell, 1)$ and we calculate
	\[t^{*\prime} = \left(\frac53, \frac43\right) \qquad x^\prime(t^{*\prime}) = \left(\frac59, \frac49\right)\]\pause

@@

\begin{frame}{Step distribution functions | Pricing N.E. Algorithm}
	\textbf{Summary:} Given $\Gm = ([2], \ell, \alpha), \alpha = (\alpha_k)_{k \in [m]}$\pause
	\begin{enumerate}
		\item Consider $\Gm^\prime = ([2], \ell, 1)$ and calculate $t^{*\prime}, x^\prime(t^{*\prime})$.\pause
		\item Calculate $\as(t^*)$ as the $\alpha$ value of the lower split in $x^\prime(t^{*\prime})$.\pause
		\item Calculate the candidate $t^* = \tfrac{t^{*\prime}}{\as(t^*)}$ and $\Pi_1(t^*), \Pi_2(t^*)$.\pause
		\item For each $\alpha_k \ne \as(t^*)$, and for each $i \in [2]$:\pause
		\begin{enumerate}[$4.1$]
			\item Consider $\Gm_k = ([2], \ell, \alpha_k)$ and calculate $B_{i, k}(t_{-i}^*)$ and tolls $t_{i, k}$ for which $\as(t_i, t_{-i}^*) = \alpha_k$, resulting in two toll ranges for $t_i$.\pause
			\item Search for $\Pi_{i, k}$, the max profit for $i$ within those $t_i$ ranges:\pause
			\begin{enumerate}[$4.2.1$]
				\item If $B_{i, k}(t_{-i}^*)$ is within the $t_i$ ranges, $\Pi_{i, k} = \Pi_i(B_{i, k}(t_{-i}^*), t_{-i}^*)$.\pause
				\item Otherwise, $\Pi_{i, k}$ is max profit among the ranges' edge tolls ($4$).
			\end{enumerate}\pause
			\item If $\Pi_i < \Pi_{i, k}$, then $\Gm$ has NO pricing Nash Equilibrium.
		\end{enumerate}\pause
		\item If all of the above checks fail, return $t^*$.
	\end{enumerate}\pause
	\textbf{Complexity:} $\mathcal{O}(m)$
\end{frame}

@@

\begin{frame}{Future work}
	\textbf{Pseudo-heterogeneous games}\pause
	\begin{itemize}
		\item Insights in fixed distribution function variation\pause
		\item What happens to flow, tolls, profit when fixed $t_{-i}$?\pause
		\item (Conjectures in the Appendix)
	\end{itemize}\pause

@@

\begin{frame}{Future work}
	\textbf{Continuous distribution functions}\pause
	\begin{itemize}
		\item No gaps!\pause
		\item Function series approximation ($m \rightarrow +\infty$)\pause
		\item Strange behaviour when $0$ is an accumulation point in $\alpha$
	\end{itemize}\pause
